Nearly $30$ years after the introduction of the first lattice-based cryptosystem~\cite{Ajtai96}, lattice-based cryptography is now in widespread deployment. Following NIST standardization of the ML-KEM key exchange mechanism and the ML-DSA and Falcon signatures, lattice-based primitives have been integrated into applications used by billions daily---including TLS, SSH, Signal and iMessage, Zoom, WireGuard and IKEv2, and MLS---while lattice-based fully homomorphic encryption sees increasing real-world deployment as well.

As lattice-based cryptography and succinct arguments have both matured, interest in \emph{proving statements about lattice-based cryptography} has grown in tandem. One important use case is building efficient versions of more complex primitives---aggregate, blind, group, multi, threshold, and ring signatures---by verifying simpler lattice-based signatures inside a SNARK~\cite{PQCRYPTO:BCOS20,AC:LyuNgu22,C:AABKT24}. Proving FHE operations, Ajtai hash commitments, and lattice-based VRFs share the same algebraic core: proving statements of the form $As = t$ over a cyclotomic ring $R_q = \integers_q[X]/(X^n + 1)$.

\parhead{The ring dimension barrier.}
Prior approaches to proving ring arithmetic with hash-based SNARKs face a fundamental tradeoff between the ring dimension $n$ and either prover cost or proof size.
Arguments that work directly over cyclotomic rings $R_q$ pay a linear proof size cost in $n$, where concretely $n \ge 512$ at the 128-bit security level.
On the other hand, the most direct approach to simulating ring arithmetic over fields---decomposing ring elements into $n$ coefficients over $\field_p$ and encoding ring multiplication as an NTT butterfly circuit~\cite{PQCRYPTO:BCOS20}---incurs $O(Mn\log n)$ constraints per inner product of $M$ ring elements, inflating the prover by a $\log n$ factor beyond the witness size.
Sumcheck-based approaches that compose FFT, Hadamard-product, and inverse-FFT sub-protocols via GKR~\cite{CCS:LiuXieZha21} match our prover cost $O(NMn)$ but achieve a lower $O(\log(NMn))$ verifier by paying only $O(\log n)$ per batched NTT instance. The innovative ring-switching approach of Huang, Mao, and Zhang~\cite{EPRINT:HuaMaoZha25} avoids both costs by recasting the statement over a Galois ring, but requires a quotient polynomial oracle $\tilde{Q}$ whose degree scales with $n$, inflating commitment costs from $O(nM)$ to $O(nNM)$ in the holographic (public-$A$) setting---a $4\times$ blowup for typical R1CS instances where $M \approx N$. Their oracles are typed over Galois rings rather than fields, and the resulting polynomial commitments still lag behind field-based counterparts.

\parhead{Our approach: a CRT-domain PIOR for ring arithmetic.}
We introduce an efficient polynomial interactive oracle reduction (PIOR) for proving $As = t$ over the fully splitting cyclotomic ring $R_p = \integers_p[X]/(X^n + 1)$. Our starting observation is that since the CRT isomorphism $R_p \cong \field_p^n$ converts ring multiplication to pointwise multiplication, we can fold the CRT matrix $\tilde{C}$ directly into the virtual polynomial of a \emph{single} sumcheck expressing the full matrix-vector product, rather than composing separate transform sub-protocols. The CRT-domain virtual polynomials $A_C$, $s_C$, $t_C$ are never materialized as committed oracles: they are defined implicitly from the coefficient-domain oracles $\tilde{A}$, $\tilde{s}$, $\tilde{t}$ and the CRT-matrix multilinear extension $\tilde{C}$, and a second sumcheck resolves the CRT-domain claims back to evaluation claims on the committed coefficient-domain oracles. We show that the CRT evaluation tables needed for the first sumcheck can be built via the NTT in $O(Mn\log n)$ time, yielding:
\begin{itemize}[nosep]
    \item \emph{Prover:} $O(NMn + Mn\log n)$, which is $O(NMn)$ whenever $N \ge \log n$.
    \item \emph{Verifier:} $O(n + \log NM)$ field operations plus $3$ oracle queries.
    \item \emph{Proof size:} $2\mu + 5\nu + 3$ elements of $\extensionfield$, i.e., $O(\log Mn)$ with no dependence on $n$.
\end{itemize}
A holographic specialization for public $A$ eliminates the second sumcheck entirely, reducing the proof size to $2(\mu + \nu) + 1$ elements (all degree-$2$ rounds) and leaving a single evaluation claim on the committed witness oracle.

Compared to the prior state of the art, this protocol removes $n$ from the proof size (versus ring-native approaches), removes a factor of $\log n$ from the prover (versus NTT-circuit unrolling), and reduces the ring dimension's contribution to the verifier from multiplicative to additive. The resulting $O(n)$ additive verifier term arises from two evaluations of $\tilde{C}$ (\cref{lem:m-eval}), which we show can each be computed in $O(n)$ field operations by exploiting the dyadic periodicity of the CRT roots. Before arriving at this CRT-domain protocol, we developed a coefficient-domain sumcheck that proves ring multiplication directly as a negacyclic circulant matrix--vector product, bypassing the CRT decomposition entirely and recovering an $O(\log n)$ verifier. However, the coefficient-domain convolution entangles the column-index and coefficient-index variables, producing folding challenges that fix the oracle evaluation points at distinct last-$\nu$ coordinates. This precludes the shared-coordinate PCS consolidation that our CRT-domain protocol enables. In our ring Plonkish SpeedySpartan example (\cref{sec:reducing-commitments}), these PCS optimizations---including selector-polynomial merging, free multi-point reduction, ring-switch packing, and interleaved BaseFold opening---save the verifier significantly more time than the single $O(n)$ CRT evaluation costs, given that $n$ typically ranges from $64$ to $512$ at the $128$-bit security level.

\parhead{Composability via the PIOR framework.}
We prove round-by-round knowledge soundness (RBR-KS) for our protocol by decomposing it into four composable PIOR building blocks---$\Pi_{\alpha,\alpha'}$, $\Pi_{\mathsf{sc1}}$, $\Pi_\xi$, $\Pi_{\mathsf{sc2}}$---whose knowledge errors are bounded independently and concatenated via the composition theorem of~\cite{BCFW25}. This decomposition, visualized as a claim-flow DAG (\cref{fig:ringmul-pior-dag}), makes the security analysis modular: each PIOR box has a clearly defined source and target relation, and new reductions can be attached to the DAG's output without revisiting the inner protocol's soundness proof. In particular, the evaluation claims produced by our ring-multiplication PIOR feed directly into further sumcheck-based reductions (packing, multi-point-to-single-point, BaseFold folding), and in principle into any other RBR-KS protocol that consumes evaluation claims. This composability means our ring-arithmetic PIOR could be used as a subroutine within a larger proof system that combines ring-arithmetic checks with other types of constraints, with the sumchecks potentially sharing challenges or folding into a unified protocol.

\parhead{Field universality: from binary to Mersenne primes.}
Our PIOR operates over any extension field $\extensionfield/\field_p$ of sufficient size, making it compatible with a wide variety of polynomial commitment schemes.
\begin{itemize}[nosep]
    \item When $p^k - 1$ has high $2$-adicity (the ``NTT-friendly'' case, which includes Goldilocks and BabyBear primes), the PCS can be instantiated with Reed--Solomon codes via FRI~\cite{BBHR18} or WHIR~\cite{EPRINT:ACFY24}, yielding the best concrete proof sizes and verification times.
    \item When instead $p^k + 1$ has high $2$-adicity (e.g., the Mersenne prime $p = 2^{31}-1$, where $p + 1 = 2^{31}$), Circle STARK polynomial commitments~\cite{EPRINT:HabLevPap24} can be used: they exploit the circle group of order $p^k + 1$ over $\field_{p^k}$ in place of the multiplicative group.
    \item For $p = 2$, the Binius PCS~\cite{EPRINT:DiaPos24} commits multilinear polynomials over binary tower fields with zero embedding overhead, and is particularly efficient for Boolean-heavy computations.
    \item For arbitrary fields where none of the above apply, random foldable linear codes (as in Basefold~\cite{basefold} or Brakedown~\cite{C:GLSTW23}) provide a universal baseline that works over any field.
\end{itemize}
Our construction thus covers the widest set of field characteristics of any ring-arithmetic IOP we are aware of. That said, the choice of base field should be driven primarily by the computation being arithmetized: the arithmetization overhead from a mismatched base field will almost always dominate the efficiency difference between polynomial commitment schemes, so choosing the right arithmetization domain trumps choosing the best PCS.

\parhead{An odd-prime tower for packing.}
To benefit from the ring-switch packing technique of Diamond and Posen~\cite{EC:DiaPos25}---which reduces the polynomial dimension by $s = \log_2[\extensionfield : \field_p]$ at the cost of an $s$-round degree-$1$ sumcheck---we require a tower of quadratic extensions $\field_p \subset \field_{p^2} \subset \cdots \subset \extensionfield$ with an explicit multilinear basis at each level.
Diamond and Posen's original construction targets the binary tower over $\field_2$.
We give a new tower construction for all odd primes $p$ (\cref{sec:odd-prime-tower}), built from iterated quadratic extensions $\tau_\iota = \tau_{\iota-1}[X_{\iota-1}]/(X_{\iota-1}^2 - X_{\iota-2})$ with a carefully chosen non-square $\alpha \in \field_p$ at the base.
Choosing $\alpha$ as small as possible (e.g., $\alpha = 3$ for $p = 257$) makes the ``soft'' multiplications (by compile-time constants) essentially free, and the Karatsuba-recursive tower multiplication achieves $k^{\log_2 3}$ hard $\field_p$-multiplications for a degree-$k$ extension---matching the binary tower's hard-multiplication count.
This tower is needed specifically for the packing step; the core ring-arithmetic PIOR itself only requires a sufficiently large extension field.

\parhead{Application beyond lattice cryptography: SWIFFT as an in-circuit hash.}
Our protocol is useful even when the circuit being proved was not designed to verify lattice cryptography. As a motivating example, one can replace the in-circuit hash function in an existing SNARK with SWIFFT~\cite{FSE:LMPR08}, the lattice-based hash whose collision resistance reduces to worst-case ideal-lattice problems.\footnote{%
    SWIFFT provides collision resistance, not random-oracle security. Following the BHP-BLAKE2s construction of~\cite{plumo}, one can compose SWIFFT with a single iteration of the BLAKE3 compression function---applied to the (short) SWIFFT output---to obtain a hash suitable for instantiating a random oracle.}
To do so, one reparameterizes SWIFFT to use the modulus $p$ of the field one already works over for the rest of the circuit, and sets the ring dimension $n$ and norm bound $\beta$ to achieve the desired RSIS security level.
For large moduli, inputs to SWIFFT need not be represented in binary: one can use a higher-radix representation, with each digit a value in $\{0, \ldots, B-1\}$ for an appropriate base $B$, reducing the number of ring elements in the RSIS instance accordingly.
This approach is particularly attractive in light of ongoing cryptanalytic concerns regarding algebraic hash functions optimized for proof systems: recent Gröbner-basis analyses~\cite{EPRINT:GKR25,EPRINT:BBLP22} have identified inaccuracies in Poseidon's original security model that lead to misestimated round requirements, while SWIFFT's security rests on the well-studied worst-case hardness of ideal lattice problems and admits no known structural attacks.
Moreover, SWIFFT is one of the fastest hash functions both in-circuit (when the field matches the SWIFFT modulus) and out-of-circuit, where NTT-based evaluation is competitive with SHA-256 throughput.

Systems like SWIRL~\cite{swirl}, which stack WHIR with LogUp to handle lookups across AIR traces of different sizes within a single proof, may provide further opportunities in this direction.
In particular, if one picks the SWIFFT modulus to be $p^t$ for $t$ large enough that RSIS over $\integers_{p^t}[X]/(X^n + 1)$ is secure for norm bound $\beta = p$, then every $\field_p$-element fed to SWIFFT already satisfies the norm constraint without any digit decomposition. Cross-subfield lookup mechanisms could then handle the interaction between $\field_p$-arithmetic in the main circuit and $\field_{p^t}$-arithmetic in the SWIFFT sub-circuit within a unified proof.

\parhead{Comparison with prior work.}
We briefly compare with the main prior approaches for proving ring arithmetic, with a detailed survey in \cref{sec:related-work}. Unrolling ring multiplication into an NTT butterfly circuit~\cite{PQCRYPTO:BCOS20,EC:BCRSVW19} contributes $O(Mn\log n)$ constraints regardless of the backend, a $\log n$ factor beyond the witness size; our prover is linear in the witness. The GKR-based sumcheck composition of~\cite{CCS:LiuXieZha21} matches our $O(NMn)$ prover but achieves $O(\log(NMn))$ verification, whereas our verifier pays $O(n + \log NM)$; as discussed above, the additive $O(n)$ is a worthwhile trade for the PCS consolidation it enables. The ring-switching approach of~\cite{EPRINT:HuaMaoZha25} requires committing to quotient polynomial oracles typed over Galois rings, inflating commitment costs and precluding the DP25 packing reduction. Galois-ring proof systems~\cite{C:JLXYY25,PKC:WeiZhaDen25} target a complementary setting---general-purpose $\integers_{2^k}$-arithmetic---and cannot directly exploit the field-tower packing that our fully-splitting-field approach enables. The FHE-oriented GKR protocol of~\cite{EPRINT:PLFT25} amortizes across CRT slots in parallel, committing in the CRT basis, whereas our commitments are in the coefficient basis and our CRT-domain polynomials remain virtual.

\parhead{Implementation: Microlotus PCS and ring Plonkish SpeedySpartan.}
We implement \textsc{Microlotus}, a multilinear polynomial commitment scheme combining Basefold, random linear codes, our odd-prime tower construction, ring-switch packing, and oracle skipping. Instantiated over the degree-$16$ tower extension of $\field_{257}$---the natural setting for SWIFFT---\textsc{Microlotus} achieves a prover over $10\times$ faster than WHIR and Basefold, a verifier over $10\times$ faster than Brakedown and $2\times$ faster than Basefold for larger polynomials, with proof sizes roughly $2\times$ those of WHIR and Basefold but far smaller than Brakedown's $O(\sqrt{n})$. These benchmarks reinforce the principle that the right choice of arithmetization domain (here, an odd-prime tower matching the computation's native modulus) dominates the PCS-level efficiency differences.

We illustrate the full SNARK pipeline with a ring Plonkish SpeedySpartan~\cite{speedyspartan} that verifies arithmetic circuit satisfiability over $R_p$. The holographic variant of our PIOR, combined with the CRT-inlining optimization that eliminates the second sumcheck, produces evaluation claims whose shared coordinates enable a chain of zero-cost or low-cost reductions: selector-polynomial merging consolidates four indexed oracles into one, multi-point reduction exploits the full Boolean-hypercube covering for free, ring-switch packing compresses dimensions by $s = \log_2[\extensionfield : \field_p]$, and interleaved BaseFold opening shares Merkle authentication paths across the two surviving claims. The final PIOR chain reduces five evaluation claims on five oracles to two packed Basefold openings with shared folding. The efficiency and cleanliness of this pipeline is a direct consequence of our CRT-domain protocol design: by ensuring that all evaluation claims share a common set of coordinates, each downstream optimization applies at full strength.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Contributions}
\label{sec:contributions}

This work makes the following contributions:
\begin{enumerate}[nosep]
    \item \textbf{A PIOR for cyclotomic ring multiplication} (\cref{sec:matrix-vector-pior}). We present a sumcheck-based PIOR for proving $As = t$ over fully splitting cyclotomic rings $\integers_p[X]/(X^n + 1)$, with $O(NMn)$ prover complexity, $O(n + \log NM)$ verifier complexity, and $O(\log Mn)$ proof size. An NTT-based evaluation-table construction eliminates the $O(Mn^2)$ bottleneck of na\"ive CRT-domain expansion. A holographic variant for public $A$ eliminates the second sumcheck, reducing proof size to $2(\mu + \nu) + 1$ with all degree-$2$ rounds and a single committed-oracle evaluation claim. We prove round-by-round knowledge soundness via a compositional PIOR decomposition. The $f$-circulant formulation also covers odd cyclotomics via the Chinese remainder transform.
    \item \textbf{Codes and proximity gaps for cyclotomic rings} (\cref{sec:ring-codes}). We show that any linear $[n, k, d]$-code over a field $\field_q$ gives a linear $[n, k, d]$-code for all cyclotomic rings $\integers_q[X]/(\Phi_m(X))$. We prove that linear codes over all finite rings $R$ satisfy a $(d/3, E)$-proximity gap in the unique decoding radius regime, where $E$ is determined by the residue fields of $R$. This yields an IOP of proximity to the ring, which in turn gives a hash-based multilinear PCS over cyclotomic rings.
    \item \textbf{An odd-prime tower construction} (\cref{sec:odd-prime-tower}). We construct a tower of quadratic extensions over any odd prime $p$, generalizing the Diamond--Posen packing technique~\cite{EC:DiaPos25} from binary to odd-characteristic fields. The tower achieves $k^{\log_2 3}$ hard $\field_p$-multiplications for degree-$k$ extensions.
    \item \textbf{\textsc{Microlotus}: a multilinear PCS implementation} (\cref{sec:microlotus}). We implement a PCS combining Basefold, random linear codes, our odd-prime tower, ring-switch packing, and oracle skipping. Over a degree-$16$ tower of $\field_{257}$, the prover is over $10\times$ faster than WHIR and Basefold for polynomials of practical size.
    \item \textbf{A ring Plonkish SpeedySpartan} (\cref{sec:reducing-commitments}). We instantiate a SpeedySpartan-style SNARK for ring arithmetic and show how the CRT-domain protocol design enables a chain of zero-cost reductions---selector merging, free multi-point reduction, ring-switch packing, and interleaved BaseFold opening---that consolidates five evaluation claims on five oracles into two packed Basefold openings with shared folding.
\end{enumerate}
