\section{Preliminaries}
\label{sec:preliminaries}

\parhead{Notation}
We write $[n]$ for the set $\{1,\dotsc,n\}$ and $\bits$ for $\{0,1\}$. We denote by $\lambda$ the security parameter, by $\negl(\lambda)$ a function negligible in $\lambda$, and by $\poly(\lambda)$ an unspecified polynomial in $\lambda$. For a finite set $S$, we write $x \sample S$ to denote sampling $x$ uniformly at random from $S$. We let $\field$ denote a finite field and $\extensionfield$ an extension field of $\field$. For a positive integer $q$, we write $\integers_q$ for the ring $\integers/q\integers$. 

An \emph{oracle} to a string or vector $\pi$ of elements from a ring $R$ is a black-box that allows querying any position $j$ and obtaining $\pi_j$. We denote it by $\oracle{\pi}$. An \emph{oracle to a polynomial} $f(\XX)$ over $R$ is a black-box that allows querying any point $\rr$ in the domain of $f$ and obtaining $f(\rr)$. We denote it by $\oracle{f}$.

\subsection{Multilinear polynomials}
\label{sec:multilinear-polynomials}

\begin{lemma}[Lagrange interpolation of multilinear polynomials]
  \label{lemma:lagrange}
  Let $f: \bits^v \to \field$ be any function. Then the following multilinear polynomial $\hat{f}$ extends $f$:
  \[
  \hat{f}(X_1,\dotsc,X_v) = \sum_{w \in \bits^v} f(w) \cdot \chi_{w}(X_1,\dotsc,X_v) 
  \]
  where, for any $w = (w_1,\dotsc,w_v)$,
  \[
  \chi_{w}(X_1,\dotsc,X_v) := \prod_{i=1}^v \bigl(X_i w_i + (1-X_i)(1-w_i)\bigr)
  \]
  \noindent The set $\{ \chi_{w} : w \in \bits^v \}$ is referred to as the set of multilinear Lagrange basis polynomials with interpolating set $\bits^v$.
\end{lemma}

We will also sometimes write $\eq(w,X)$ instead of $\chi_{w}(X)$ to denote the same polynomial.

\begin{lemma}
  \label{lemma:partial-sumcheck}
Let $p : \field^\ell \to \field$ be a multilinear polynomial. Then for any $0 \le i \le \ell$, any $(\alpha_1, \dots, \alpha_i) \in \field^i$, and any $x' \in \{0,1\}^{\ell - i}$,
\[
p(\alpha_1, \dots, \alpha_i, x') = \sum_{y \in \{0,1\}^i} \eq((\alpha_1, \dots, \alpha_i), y) \cdot p(y, x')
\]
\end{lemma}
Setting $i = 1$, the identity in \cref{lemma:partial-sumcheck} simplifies to
\[
p(\alpha_1, x') = (1 - \alpha_1) \cdot p(0, x') + \alpha_1 \cdot p(1, x')
\]
\noindent Setting $i = \ell$, we obtain the multilinear extension formula
\[
p(\alpha_1, \dots, \alpha_\ell) = \sum_{y \in \{0,1\}^\ell} \eq((\alpha_1, \dots, \alpha_\ell), y) \cdot p(y)
\]


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Cyclotomic Rings}
\label{sec:rings}

% In this work we will assume all rings $\ring$ to be unital and commutative. Though we will not cover ring theory in any depth, we briefly remind the reader of a few key concepts we make use of in this paper:
% \begin{itemize}[nosep]
%   \item \emph{Maximal ideal}: an ideal $\maximalideal$ that is maximal (with respect to set inclusion) amongst all proper ideals. In other words, no ideal $\ideal$ is contained between $\maximalideal$ and $\ring$.
%   \item \emph{Local ring}: a ring with a unique maximal ideal. Here the sum of any two non-units is a unit, for any $r \in \ring$ either $r$ or $1-r$ is a unit, and if any finite sum is a unit, then it has a term that is a unit.
%   \item \emph{Semilocal ring}: a ring with finitely many maximal ideals. 
%   \item \emph{Residue field}: for any maximal ideal $\maximalideal$, the residue field is the quotient ring $\ring/\maximalideal$, which is a field.
%   \item \emph{Reduced ring}: a ring with no non-zero nilpotent elements. Equivalently, if $r^2 = 0$, then we must have $r = 0 \in \ring$.
% \end{itemize}

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \subsubsection{Cyclotomic rings}
% \label{sec:cyclotomic-rings}

Let $\zeta = \zeta_m$ be a primitive $m$th root of unity (i.e., $\zeta^m = 1$, but $\zeta^i \neq 1$ for $i \in [m-1]$). For any positive integer $m$, the $m$th cyclotomic polynomial $\Phi_m$ is the minimal polynomial of $\zeta$ over the integers:
% the unique, irreducible polynomial with integer coefficients that is a divisor of $X^{m}-1$ and not $X^{k}-1$ for any $1 \leq k < m$. It can be written
\[
  \Phi_m(X) = \prod_{\substack{k \in [m], \\ \gcd(k,m) = 1}} \left(X - \zeta^k\right)
\]
where we observe its conjugates are the other primitive $m$th roots of unity. There is a natural isomorphism between $\integers[\zeta]$ and $\ring = \integers[X]/\Phi_m(X)$ given by $\zeta \mapsto X$. Let $n = \phi(m)$. Then $\bm{p} = (\zeta^i)_{i=0}^{n-1}$ defines a $\integers$-basis for $\ring$, called the \emph{power basis}. We note that any $\integers$-basis for $\ring$ is a $\integers_q$-basis for $\ring_q = \integers_q[X]/\Phi_m(X)$.

Another important $\integers_q$-basis for $\ring_q$ is the Chinese remainder transform basis, or \emph{CRT basis}, where ring multiplication simplifies to computing the Hadamard product of the $\integers_{q}^n$ vectors. The Chinese remainder transform can be seen as a specialization of the discrete Fourier transform to cyclotomic rings, and is also called the number theoretic transform (NTT). By virtue of being the rings of integers of algebraic number fields, cyclotomic rings are Dedekind, which means ideals can be factored into a product of prime ideals. A particular case of interest is $q \equiv 1\ \mod m$, which corresponds to the ``fully splitting'' case. Notice that $\integers_q$ contains a primitive $m$th root of unity $\omega_m$. Then the principal ideal $\langle q \rangle$ in $\ring$ can be written as the product of the prime ideals $\primeideal_{i} = \langle q \rangle + \langle \zeta_m - \omega_{m}^i \rangle$ for $i \in \integers_{m}^{*}$, i.e., it splits completely into $n$ ideals of norm $q$ each. Each quotient ring $\ring/\primeideal_{i}$ is isomorphic to the field $\integers_q$ via the map $\zeta_m \mapsto \omega_{m}^i$. Then by the Chinese remainder theorem
\[
\ring/\ideal \cong \prod_i \ring/\primeideal_{i} \cong (\integers_{q})^n
\]
The CRT basis $\bm{c}$ is then defined such that $c_i \equiv 1 \bmod \primeideal_{i}$ and $c_i \equiv 0 \bmod \primeideal_j$ for all $j \neq i$.

The Chinese remainder transform and its inverse convert between the power and CRT bases. The linear transform, given as a matrix $\crt$ below, has the identity $\bm{p}^\intercal = \bm{c}^\intercal \cdot \crt$.

\begin{definition}[\cite{EC:LyuPeiReg13}]
  \label{def:crt}
Let $m$ be a prime power and let $R$ denote any commutative ring containing some element $\omega_m$ of multiplicative order $m$, i.e., a primitive $m$th root of unity.
\begin{itemize}[nosep]
  \item The \emph{discrete Fourier transform} $\mathsf{DFT}_m$ over $R$ is the $\mathbb{Z}_m \times \mathbb{Z}_m$ matrix whose $(i,j)$th entry is $\omega_m^{ij}$.
  \item The \emph{Chinese remainder transform} $\mathsf{CRT}_m$ over $R$ is the (square) submatrix of $\mathsf{DFT}_m$ obtained by restricting to the rows indexed by $\mathbb{Z}_m^*$ and the columns indexed by $[\varphi(m)]$.
\end{itemize}
\end{definition}

Applying $\mathsf{DFT}_m$ corresponds to evaluating a polynomial of degree less than $m$ over $\ring[Y]$ (represented by the vector of coefficients in the natural order) at all the $m$th roots of unity. Similarly, $\mathsf{CRT}_m$ corresponds to evaluating a polynomial of degree less than $\phi(m)$ at the primitive $m$th roots of unity $\omega_{m}^i$ for $i \in \integers_{m}^{*}$.

% \jnote{Add definition of Gram matrix and unimodular matrix.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \subsection{Circulant matrices}
% \label{sec:circulant-matrices}

% Multiplication in the power basis of a power-of-two cyclotomic ring can be expressed as a negacyclic convolution or skew-circulant matrix-vector multiplication. We consider the following generalization of circulant matrices:

% \begin{definition}[$f$-circulant matrix]
%   \label{def:f-circulantmatrix}
%     Given a vector $a$, the $f$-circulant matrix $Z_f(a)$ is defined as
%     \[
%     Z_f(a) =
%     \begin{bmatrix*}
%         a_0 &  f\cdot a_{n-1}  & \dotsm & f\cdot a_1 \\
%         a_1 & a_0 & \ddots & \vdots \\
%         \vdots & \vdots & \ddots & f\cdot a_{n-1} \\
%         a_{n-1}  & a_{n-2} & \dotsm & a_0
%     \end{bmatrix*}
%     \]
%     For $f=1$ this corresponds to circulant matrices, and $f=-1$ corresponds to skew-circulant matrices.
% \end{definition}

% Since $Z_1(a) = F_{n}^{-1} \cdot \mathsf{diag}(F_n a) \cdot F_n$, where $F_n$ is the DFT matrix, circulant matrix-vector multiplication can be computed in $O(n \log n)$ time using $3$ FFTs. A method given in ~\cite{Rosowski20} runs in $O(n \log n)$ without FFTs, but still requires primitive $n$th roots of unity. In general,
% \begin{align*}
% Z_{f}(a) = \diag(\eta) \cdot F^{-1}_{n} \cdot \diag(F \cdot \diag(\eta) \cdot a) \cdot F \cdot \diag(\eta)
% \end{align*}
% for $\eta = [1, f^\frac{1}{n}, \dotsc, f^{\frac{n-1}{n}}]$~\cite{ClinePlemmonsWorm74}. In other words we compute can compute ring multiplication in the power basis as $Z_{-1}(a) \cdot b = \eta \circ F^{-1}_{n}(F_{n}(\eta \circ a) \circ F_{n}(\eta \circ b))$.

% For $n$ a power-of-two, this gives us another way to write the CRT matrix: $C_{2n} = F_n \cdot \diag(\eta)$. This can be confirmed by observing the $(i,j)$th entry of such a matrix is $\omega_{2n}^{2ij} \cdot \omega_{2n}^{j}$; for first term we use the fact a primitive $n$th root $\omega_n$ can be written as $\omega_{2n}^{2}$ for some primitive $2n$th root $\omega_{2n}$, and the second follows from the related fact the $n$th roots of $-1$ are the primitive $2n$th roots of unity. Simplifying, this gives $\omega_{2n}^{2ij+j} = \omega_{2n}^{(2i+1)\cdot j} = \omega_{2n}^{\integers_{2n}^{*}(i) \cdot j}$ as expected.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Indexed relations}
\label{sec:indexed-relations}

An \emph{indexed relation} is a set $\relation \subseteq \{0,1\}^* \times \{0,1\}^* \times \{0,1\}^*$ of triples $(\index, \instance, \witness)$, where $\index$ is called the \emph{index}, $\instance$ is the \emph{instance} (or \emph{statement}), and $\witness$ is the \emph{witness}. The corresponding \emph{indexed language} is $\langrel = \{(\index, \instance) : \exists\, \witness \text{ s.t.\ } (\index, \instance, \witness) \in \relation\}$.% We say $\relation$ is an NP relation if membership in $\relation$ can be decided in time $\poly(|\index| + |\instance| + |\witness|)$.
An indexed relation separates the \emph{public structure} of a problem (the index) from the \emph{per-instance} data. In the holographic setting, an indexer $\indexer$ preprocesses $\index$ into an oracle that the verifier can query during the protocol. For example, in our ring-arithmetic PIOP (\cref{sec:ring-piop}), the index encodes the public matrix $A$ (and possibly $t$), the instance is empty or encodes public parameters, and the witness is the secret vector $s$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Polynomial interactive oracle proofs}
\label{sec:piop}

We recall the definition of polynomial interactive oracle proofs (polynomial IOPs, or PIOPs)~\cite{BCRSVW19,CHMMVW20,ACFY24}. A PIOP is an interactive proof system in which the prover's messages are polynomials that the verifier can query at points of its choosing, rather than reading in their entirety.

\begin{definition}[Polynomial IOP {\cite[Section~3.1]{ACFY24}}]
\label{def:poly-iop}
A \emph{$\mu$-round polynomial IOP} (poly-IOP) for a relation $\relation$ is a pair of interactive algorithms $\Pi = (\prover, \verifier)$ where:
\begin{itemize}
    \item At the start of the protocol, $\prover$ receives $(\instance, \witness)$ and $\verifier$ receives $\instance$.
    \item The protocol proceeds in $\mu$ rounds. In round $i \in [\mu]$, there is a prescribed list of $n_i$ individual degree bounds $(d_{i,1}, \ldots, d_{i,n_i})$ with $d_{i,j} \in \mathbb{N}$. The prover sends $n_i$ polynomials by specifying their coefficients: the $j$-th polynomial $\hat{f}_{i,j} \in \extensionfield^{\leq d_{i,j}}[X]$ has degree at most $d_{i,j}$. The verifier then replies with a uniformly random challenge $c_i \sample \chalspace_i$. After round $\mu$, the prover sends a final list of polynomials.
    \item The verifier has \emph{oracle access} to each prover polynomial: it may query $\hat{f}_{i,j}$ at any point $z \in \extensionfield$, receiving the evaluation $\hat{f}_{i,j}(z)$, rather than reading the polynomial in its entirety.
    \item After the interaction, the verifier outputs $\mathsf{accept}$ or $\mathsf{reject}$.
\end{itemize}
More generally, in a \emph{multilinear} poly-IOP, each prover oracle is a multilinear polynomial $\hat{f}_{i,j} \in \extensionfield^{\multilin}[X_1, \ldots, X_{d_{i,j}}]$, and the verifier may query it at any point in $\extensionfield^{d_{i,j}}$. We write $\langle \prover(\witness), \verifier \rangle(\instance)$ for the output of the verifier after interacting with $\prover$ on common input $\instance$.
\end{definition}

An \emph{$i$-round partial transcript} is a tuple $\tau_i = (\instance, f_1, c_1, \ldots, f_i, c_i)$, where $f_i$ denotes the list of prover polynomials in round $i$; a \emph{$0$-round partial transcript} is simply $(\instance)$. A \emph{complete transcript} appends the final prover polynomials to a $\mu$-round partial transcript; it is \emph{accepting} if the verifier outputs $\mathsf{accept}$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\parhead{Completeness and soundness}

PIOP completeness is defined as follows.

\begin{definition}[Completeness]
\label{def:completeness}
A poly-IOP $\Pi = (\prover, \verifier)$ for a relation $\relation$ has \emph{(perfect) completeness} if for all $(\instance, \witness) \in \relation$,
\[
    \Pr\bigl[\langle \prover(\witness), \verifier \rangle(\instance) = \mathsf{accept}\bigr] = 1,
\]
where the probability is over the verifier's random challenges.
\end{definition}

We omit the standard soundness definition as we prove the stronger round-by-round \emph{knowledge} soundness for our protocols. We use the generalized formulation from Block, Garreta, Tiwari, and Zaj\k{a}c~\cite{BGTZ24}.

\begin{definition}[Round-by-round knowledge soundness {\cite[Definition~4.2]{BGTZ24}}]
\label{def:rbr-knowledge}
A poly-IOP $\Pi = (\prover, \verifier)$ for a relation $\relation$ has \emph{round-by-round knowledge soundness with errors $(\varepsilon_1, \ldots, \varepsilon_\mu)$} if there exists a doomed set $\doomedset$ and a polynomial-time algorithm $\extractor$ (the \emph{extractor}) such that:
\begin{enumerate}
    \item \textbf{All statements start doomed.} For \emph{all} inputs $\instance$ (including $\instance \in \langrel$), $(\instance) \in \doomedset$.
    \item \textbf{Doomed transcripts reject.} For any $\mu$-round partial transcript $(\instance, \tau) \in \doomedset$ and any final prover polynomials, the verifier rejects.
    \item \textbf{Escaping doom yields a witness.} For any $(i-1)$-round partial transcript $(\instance, \tau) \in \doomedset$ (where $i \in [\mu]$) and any list of prover polynomials $f_i$ in round $i$: if
    \[
        \Pr_{c \sample \chalspace_i}\bigl[(\instance, \tau, f_i, c) \notin \doomedset\bigr] > \varepsilon_i,
    \]
    then $\extractor(\instance, \tau, f_i)$ outputs a witness $\witness$ such that $(\instance, \witness) \in \relation$.
\end{enumerate}
\end{definition}

The key distinction from plain RBR soundness is condition~(1): the doomed set is defined for all inputs, not only false statements. This enables extraction --- if a (possibly honest) prover can make the transcript escape doom with probability exceeding $\varepsilon_i$, the extractor must be able to recover a valid witness from the partial transcript alone.

\begin{remark}
RBR knowledge soundness implies \emph{round-by-round soundness} (the weaker notion where only false statements start doomed, and no extractor is required) with the same errors~\cite[Remark~4.4]{BGTZ24}. It also implies $(k_1, \ldots, k_\mu)$-special soundness with $k_i = \lceil |\chalspace_i| \varepsilon_i \rceil + 1$~\cite[Theorem~1.2]{BGTZ24}.
\end{remark}

\begin{remark}[Compilation to non-interactive arguments]
Since every PIOP is an IOP (where prover oracles are constrained to be polynomials), any PIOP with RBR (knowledge) soundness can be compiled into a non-interactive argument via the Fiat-Shamir transformation~\cite{BCS16} that is adaptively (knowledge) sound in both the random oracle model~\cite{BCS16} and the quantum random oracle model~\cite{CMS19}.
\end{remark}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \subsection{Projective spaces}
% \label{sec:projective-spaces}

% \begin{definition}[Projective space]
%   \label{def:projective-space}
%   We define the projective space $\projective^{n-1}(V)$...
% \end{definition}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Sumcheck}
\label{sec:sumcheck}

Variations of the sumcheck protocol~\cite{JACM:LFKN92} have become central to the construction of many of today's state-of-the-art PIOPs. Here we consider sumcheck instances over a field $\field$ to be claims of the form $(\sigma, g, v, d)$, where $g$ is a $v$-variate polynomial of degree at most $d$ in each variable, and $\sigma$ is its claimed sum over the $v$-dimensional hypercube:
\[
 \sigma = \sum_{x_1,\dotsc,x_v \in \bits^v} g(x_1,\dotsc,x_v)
\]
The sumcheck protocol gives a reduction of this claim to checking the evaluation of $g$ at a random point $g(\alpha_1,\dotsc,\alpha_v) = \sigma_v$. For self-containedness, we have included in~\cref{app:sumcheck} an optimized sumcheck protocol (\cref{fig:sumcheck}) and a linear time and space prover algorithm for it (\cref{fig:linear-sc}) from~\cite{BagadDaoDombThaler25}.

% %-------------------------------------------------------------------------------
% \subsubsection{Sumcheck in linear time}



% \pnote{replace sumcheck for partial evaluation with lemma describing stepwise behavior}

% \subsection{BaseFold and Random Foldable Codes}


% \subsection{Binius}
% \hnote{I will shorten this significantly }
% Diamond and Posen introduce \textit{ring-switching} in [DP24], a sumcheck-based technique 
% for reducing openings of tiny field multilinears at extension field points to openings of 
% packed multilinears entirely over an extension field. Ring-switching is generic to 
% power-of-two degree extension fields. [Check the wording on this... -Hunter].

% The Binius family of SNARKs utilize ring-switching as a component of FRI-Binius,
% a PCS for proving tiny field multilinear openings at large field points over binary fields. 
% FRI-Binius combines ring-switching with a variant of Basefold over characteristic 2.
% Instantiating Basefold in characteristic 2 is non-trivial because binary fields lack
% the power-of-two order multiplicative subgroups necessary for standard FRI proofs in which 
% folding is done over a multiplicative univariate monomial basis. The observation behind FRI-Binius
% that makes binary basefold possible is that if codewords in FRI are folded over an additive basis, 
% not a multiplicative one, it preserves the multilinear evaluation property 
% necessary for Basefold in characteristic 2. This is why the Additive-NTT is chosen as 
% the encoding scheme in FRI-Binius. [Check if this explanation makes sense to other 
% people. This explanation is in reference to Section 1.4 of DP24 which discusses the issue
% with the multiplicative NTT for Basefold in characteristic 2 ... -Hunter]

% In characteristic $257$ this issue is not present, as $\mathbb{Z}_{257}$ is a smooth prime-field, 
% containing the multiplicative subgroups necessary for running FRI with the multiplicative NTT or
% Random Foldable Codes. This makes it more straightforward to apply ring switching to Basefold 
% in characteristic $257$.

% While the Binius techniques are often thought of as being exclusive to binary fields, protocols 
% involving packing apply beyond this setting. Though, the choice of binary fields in Binius is 
% not arbirary, and contains a certain advantage. Modern processors natively operate on 
% binary data, usually in the form of $64$-bit processor words. Directly arithmetizing over binary 
% fields can enable more efficient arithmetizations when the target application begins in this setting.
% For example, when arithmetizing signed integer multiplication.

% As well, converting CPU native data types to in-memory prime-field representations is often
% wasteful, especially if those data types are small, such as bytes. A byte needs only $8$ bits
% to represent in memory, where as the prime field may require $32$ or $64$ bits. This concept
% of is refered to as \textit{embedding overhead} in the Binius termonology. Additionally, 
% if those bytes refer to small prime-fields, non-native arithmetic of a small field will need
% to be encoded into the arithmetization field. This has the consequence of adding extra gates
% to the overall circuit than would be necessary if directly arithmetized over the target field.

% Circuits of application defined over tiny fields benefit significantly from direct arithmetization 
% over their target field for this reason. In our case, lattice-base cryptographic primitives are 
% often defined over polynomial rings with coefficients in a tiny base field, such as 
% $\mathbb{Z}_{257}$. This is true for the SWIFFT compresion function, and LM08 one-time signatures.
% Our implementation \textit{microlotus} implements ring-switching for basefold in characteristic 
% $257$, enabling circuits defined directly over $\mathbb{Z}_{257}$. More is discussed on the 
% concrete performance gains of this technique in the implementation section.

% For certain applications, such as primitives in lattice-based cryptography, we may be 
% interested in arithmetizing circuits over a \textit{tiny field}. Diamond and Posen refer to 
% \textit{tiny field polynomials} as those having more coefficients than elements in 
% their defining field. Instantiating hash-based arguments regarding such polynomials 
% will encounter the \textit{trace-length barrier}, which refers to when the witness of an 
% argument is larger than the size of the field. This prevents the use of codes such as 
% Reed Solomon, which require an alphabet at least as large as the block size of the code. 

% The suite of protocols behind \textit{Binius} have emerged as feasible and performant 
% solutions to this trace-length barrier. Such solutions make use of \textit{packing} and 
% \textit{ring-switching}, which enable hash-based commitments of tiny-field 
% polynomials that achieve significant performance gains by reducing prover work. 
% Although Binius SNARKs are generally thought of as being specific to binary 
% fields, many of the techniques, such as \textit{ring-switching} apply generally to 
% tiny field polynomials. 
% [How in depth should be describe the FRI-Binius protocol? Is it sufficient to explain the 
% feasibility of our construction, or should we described the algorithm in depths? Also,
% we should probably cite the additive NTT - Hunter].