¥subsection{A faster verifier, larger proof size variant}
¥label{sec:logger-verifier}

The $O(n)$ verifier cost in‾¥cref{prot:ringmul} arises entirely from the direct evaluation of $¥tilde{C}(¥alpha', ¥gamma)$ in the final check.
We now describe a variant that reduces this to $O(¥nu^2)$ by delegating the evaluation to the prover via a sequence of sumchecks, at the cost of increasing the proof size from $O(¥mu + ¥nu)$ to $O(¥mu + ¥nu^2)$ extension field elements.

The starting point is the factored representation from the proof of‾¥cref{lem:m-eval}.
Recall that $¥tilde{C}(¥alpha', ¥gamma) = ¥sum_{k ¥in ¥bits^¥nu} ¥meq(¥alpha', k) ¥cdot g(¥omega^{2k+1})$ where $g(¥omega^{2k+1}) = ¥prod_{j=0}^{¥nu-1} f_j(k)$ and each $f_j(k) = (1-¥gamma_j) + ¥gamma_j ¥omega^{(2k+1)2^j}$ has period $2^{¥nu-j}$ in $k$.
Define partial products accumulating factors from highest to lowest period:
¥begin{equation}¥label{eq:partial-product-butterfly}
    h^{(0)} ¥;:=¥; 1, ¥qquad h^{(r)}(k) ¥;:=¥; ¥prod_{j=¥nu-r}^{¥nu-1} f_j(k) ¥quad ¥text{for } r = 1, ¥ldots, ¥nu,
¥end{equation}
so that $h^{(r)}$ has period $2^{r}$ in $k$ and $h^{(¥nu)}(k) = g(¥omega^{2k+1})$.
The recurrence $h^{(r)}(k) = h^{(r-1)}(k) ¥cdot f_{¥nu-r}(k)$ relates consecutive layers, and the target evaluation is
¥begin{equation}¥label{eq:target-butterfly}
    ¥tilde{C}(¥alpha', ¥gamma) ¥;=¥; ¥sum_{k ¥in ¥bits^¥nu} ¥meq(¥alpha', k) ¥cdot h^{(¥nu)}(k).
¥end{equation}

¥parhead{Protocol modification} In‾¥cref{prot:ringmul}, Step‾7 is replaced as follows.
The prover sends $e_C$ claimed to equal $¥tilde{C}(¥alpha', ¥gamma)$, and the verifier checks $¥sigma'' = e_C ¥cdot (e'_t + ¥xi ¥cdot e'_A + ¥xi^2 ¥cdot e'_s)$ using the oracle queries.
It remains to verify the claim $e_C = ¥tilde{C}(¥alpha', ¥gamma)$, which is done by the following sequence of sumchecks that ``unwind'' the partial products‾¥eqref{eq:partial-product-butterfly}.

¥parhead{Layer-$r$ reduction (for $r = ¥nu, ¥nu-1, ¥ldots, 1$)} At the start of layer‾$r$, the verifier holds a claim
¥[
    e^{(r)} ¥;=¥; ¥sum_{k ¥in ¥bits^r} ¥meq(v^{(r)}, k) ¥cdot h^{(r)}(k)
¥]
for some $v^{(r)} ¥in ¥extensionfield^r$.
For $r = ¥nu$, this is initialized with $e^{(¥nu)} := e_C$ and $v^{(¥nu)} := ¥alpha'$.
Expanding the recurrence $h^{(r)}(k) = h^{(r-1)}(k) ¥cdot f_{¥nu-r}(k)$, the prover and verifier run a sumcheck for
¥[
    e^{(r)} ¥;=¥; ¥sum_{k ¥in ¥bits^r} ¥meq(v^{(r)}, k) ¥cdot h^{(r-1)}(k) ¥cdot f_{¥nu-r}(k).
¥]
This sumcheck has $r$ variables.
Naively, the integrand is a product of three multilinear polynomials ($¥meq$, $¥widetilde{h^{(r-1)}}$, $¥widetilde{f_{¥nu-r}}$), which would give individual degree‾$3$.
However, the multilinear extension of $f_{¥nu-r}$ has a product-of-univariates structure:
¥[
    ¥widetilde{f_{¥nu-r}}(k) ¥;=¥; (1-¥gamma_{¥nu-r}) + ¥gamma_{¥nu-r}¥, ¥omega^{2^{¥nu-r}} ¥prod_{i=0}^{r-1}¥Big[1 + k_i¥big(¥omega^{2^{¥nu-r+i+1}} - 1¥big)¥Big],
¥]
and can therefore be absorbed into the $¥meq$ weights.
Define $w^{(r)}(k) := ¥meq(v^{(r)}, k) ¥cdot ¥widetilde{f_{¥nu-r}}(k)$; since each factor of $¥widetilde{f_{¥nu-r}}$ depends on a single variable‾$k_i$, the function $w^{(r)}$ is multilinear in‾$k$ (the product of two univariate-in-$k_i$ terms remains degree‾$1$ in‾$k_i$).
The sumcheck is then over the degree-$2$ integrand $w^{(r)}(k) ¥cdot ¥widetilde{h^{(r-1)}}(k)$, requiring $3$ extension field elements per round, $3r$ per layer.

After sumcheck challenges $u^{(r)} ¥in ¥extensionfield^r$, the claim reduces to
¥[
    ¥sigma^{(r)} ¥;=¥; w^{(r)}(u^{(r)}) ¥cdot ¥widetilde{h^{(r-1)}}(u^{(r)}).
¥]
The verifier evaluates $w^{(r)}(u^{(r)}) = ¥meq(v^{(r)}, u^{(r)}) ¥cdot ¥widetilde{f_{¥nu-r}}(u^{(r)})$ in $O(r)$ time using the product formulas above.
The prover sends $e^{(r-1)}$ claimed to equal $¥widetilde{h^{(r-1)}}(u^{(r)})$, and the verifier checks $¥sigma^{(r)} = w^{(r)}(u^{(r)}) ¥cdot e^{(r-1)}$.
The next layer begins with the claim $e^{(r-1)} = ¥sum_{k ¥in ¥bits^{r-1}} ¥meq(v^{(r-1)}, k) ¥cdot h^{(r-1)}(k)$ for appropriate $v^{(r-1)}$ derived from $u^{(r)}$.

At the base case $r = 0$, we have $h^{(0)} = 1$ and the claim $e^{(0)} = 1$ is checked directly.

¥parhead{Complexity} The total transcript for the butterfly delegation is $¥sum_{r=1}^{¥nu} (3r + 1) = ¥frac{3¥nu(¥nu+1)}{2} + ¥nu = O(¥nu^2)$ extension field elements (the $+1$ per layer accounts for the prover's claim $e^{(r-1)}$).
Combined with the main protocol's $3¥mu + 3¥nu + 2$ elements (from the two sumchecks and the claims $e_t, e_A$) plus the additional claim $e_C$, the total proof size is $3¥mu + 3¥nu + 3 + O(¥nu^2) = O(¥mu + ¥nu^2)$.
The verifier's cost is $¥sum_{r=1}^{¥nu} O(r) = O(¥nu^2)$ for the butterfly checks plus $O(¥tau + ¥mu)$ for the main sumchecks, giving $O(¥nu^2 + ¥tau + ¥mu) = O(¥log^2 n + ¥log NM)$.
The prover's asymptotic cost is unchanged at $O(NMn)$: the sumcheck tables $[h^{(r)}(k)]_{k ¥in ¥bits^r}$ for $r = 1, ¥ldots, ¥nu$ are byproducts of the iterative algorithm in‾¥cref{lem:m-eval} and require no additional work to produce.